---
title: "Data-Wrangling"
output: html_document
---

```{r setup, include=FALSE}
# DON'T TOUCH ANYTHING IN THIS BOX
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Activity 1 - Load the tidyverse

In the code chunk below, write and run the code to load the tidyverse.

```{r T01}
# hint: library(tidyverse)
library(tidyverse)
```

## Activity 2 - Load in the data

Now we have the tidyverse loaded, edit the below code to load in the three data files, `p_screen.csv`, `QHC_responses.csv`, and `QHC_scoring.csv` and then run the code (make sure you have loaded the tidyverse). You  need to replace the `NULL` values with the code used to load each of the data files. Ensure that the names of the csv files are exactly as they were when you downloaded them from Moodle.

You should use `read_csv()` to load the files and NOT `read.csv()` as it will change the names of your variables and you won't get the correct answers for the rest of the tasks!

```{r T02}
# hint: name <- read_csv("file_name.csv")
screening <- read_csv('p_screen.csv')
responses <- read_csv('QHC_responses.csv')
scoring <- read_csv('QHC_scoring.csv')
```

## View the data

1. It is always a good idea to familiarise yourself with the layout of the data that you have just loaded in. You can do this through using `glimpse()` or `View()` in the Console windown, but you must never put these functions in your assignment file.


## Activity 3 - Oldest Participant

Replace the `NULL` in the `T3` code chunk with the Participant ID of the oldest participant. Store this single value in `oldest_participant` (e.g. `oldest_participant <- 999`.

```{r T03}
# hint: look at your data, who is oldest?
oldest_participant <- 3
```

## Activity 4 - Arranging D-SPAN

Replace the `NULL` in the `T4` code chunk with code that arranges participants' D-SPAN performance from highest to lowest using the appropriate one-table dplyr (i.e., Wickham) verb. Store the output in `cogtest_sort`. (e.g. `cogtest_sort <- verb(data, argument)`)

```{r T04}
# hint: arrange your screening data
cogtest_sort <- arrange(screening, desc(DSPAN))

```

## Activity 5 - Foreign Language Speakers

Replace the `NULL` in each of the two lines of code chunk `T5`, so that `descriptives` has a column called `n` that shows the number of participants that speak a foreign language and number of participants that do not speak a foreign language, and another column called `median_age` that shows the median age for those two groups. If you have done this correctly, `descriptives` should have 3 columns and 2 rows of data, not including the header row.

```{r T05}
# hint: First need to group_by() foreign language
screen_groups <- group_by(screening,FLANG)

# hint: second need to summarise(). Pay attention to specific column names given.

descriptives <- screen_groups %>% summarise (n=n(), median_age= median(AGE))

```

## Activity 6 - Creating Percentage MOCA scores

Replace the `NULL` in the `T6` code chunk with code using one of the dplyr verbs to add a new column called `MOCA_Perc` to the dataframe `screening` In this new column should be the `MOCA` scores coverted to percentages. The maximum achievable score on MOCA is `30` and percentages are calculated as `(participant score / max score) * 100`. Store this output in `screening`.

```{r T06}
# hint: mutate() something using MOCA and the percentage formula
screening_mutate <- mutate(screening,MOCA_Perc = (MOCA / 30) * 100)
```

### Activity 7 - Remove the MOCA column

Now that we have our MoCA score expressed as a percentage `MOCA_Perc` we no longer need the raw scores held in `MOCA`. Replace the `NULL` in the `T7` code chunk using a one-table dplyr verb to keep all the columns of `screening`, with the same order, but without the `MOCA` column. Store this output in `screening`.

```{r T07}
# hint: select your columns
screening_reduced <- select(screening_mutate, -MOCA)
```


The next set of tasks focus on merging two tables. 

You suspect that the older adults with musical experience might report more finely-tuned hearing abilities than those without musical experience. You therefore decide to check whether this trend exists in your data. You measured participant's self reported hearing abilties using the Better Hearing Institute Quick Hearing Check Questionnaire. In this questionnaire participants rated the extent to which they agree or disagree with a list of statements (e.g. 'I have a problem hearing over the telephone') using a 5 point Likert scale (Strongly Disagree, Slightly Disagree, Neutral, Slightly Agree, Strongly Agree). 

Each participant's response to each question is contained in the `responses` dataframe in your environment. Each response type is worth a certain number of points (e.g. Strongly Disagree = 0, Strongly Agree = 5), and the scoring key is contained in the `scoring` dataframe. A score for each participant is calculated by totalling up the number of points across all the questions to derive at an overall score. The lower the overall score, the better the participant's self-reported hearing ability.  

In order to score the questionnaire we first need to perform a couple of steps.

### Activity 8 - Gather the Responses together

Replace the `NULL` in the `T8` code chunk using code to gather the responses to all the questions of the QHC from wide format to tidy/long format. Name the first column `Question` and the second column `RESPONSE`. Store this output in `responses_long`.  

```{r T08}
# hint: gather the question columns (Q1:Q15) in responses
responses_long <- gather(responses,key=Question, value=RESPONSE, Q1:Q15)
```

### Activity 9 - Joining the data

Now we need to join the number of points for each response in `scoring` to the participants' responses in `responses_long`. 

Replace the `NULL` in the `T9` code chunk using `inner_join()` to combine `responses_long` and `scoring` into a new variable called `responses_points`. 

```{r T09}
# hint: join them by the column common to both scoring and responses_long
responses_points <- inner_join(responses_long, scoring, by = 'RESPONSE')
```

### Activity 10  - Working the Pipes

Below we have given you a code chunk with 5 lines of code. The code takes the data in its current long format and then creates a QHC score for each participant, before calculating a mean QHC score for the two groups of participants - those that play musical intruments and those that don't - and stores it in a variable called `musical_means`.  

```{r, eval = FALSE}
participant_groups <- group_by(responses_points, ID)
participant_scores <- summarise(participant_groups, Total_QHC = sum(SCORE))

participant_screening <- inner_join(participant_scores, screening, "ID")

screening_groups_new <- group_by(participant_screening, MUSICAL)
musical_means <- summarise(screening_groups_new, mean_score = mean(Total_QHC))

```

Replace the `NULL` in the `T10` code chunk with the following code converted into a fuctioning pipeline using pipes. Put each function on a new line one under the other. This pipeline should result in the mean QHC values of musical and non-musical people stored in `musical_means` which should be made of two rows by two columns.  

* hint: function1 %>% function2

```{r T10}
# hint: in pipes, the output of the previous function is the input of the subsequent function.
musical_means <- group_by(responses_points, ID) %>%
  summarise(Total_QHC = sum(SCORE)) %>%
  inner_join(screening, "ID") %>%
  group_by(MUSICAL) %>%
  summarise(mean_score = mean(Total_QHC))

```

### Activity 11 - Difference in Musical Means

Finally, replace the `NULL` in the `T11` code chunk with a single value, to two decimal places, that is the value of how much higher the QHC score of people who play music is compared to people who don't play music (e.g. 2.93)

```{r T11}
# hint: look in musical means and enter the difference between the two means.
QHC_diff <- 1.53
```

# Research Methods Questions

The second half of this assessment will focus on questions about research methods and statistics. This will help you consolidate what you have learned from the lectures and the reading. To complete these questions you will replace the `NULL` with the number of the correct option.

## Activity 12

In an experimental design, the independent variable is:

1. The measured outcome variable
2. The variable the researcher manipulates
3. An extra variable that has a hidden effect on your experimental results.

```{r T12}
answer_t12 <- 2
```

## Activity 13

A researcher conducts a study and pre-registers their hypotheses in advance of data collection. They have specific directional hypotheses for what they expect to find. Is this study:

1. Exploratory
2. Confirmatory

```{r T13}
answer_t13 <- 2
```

## Activity 14

Publication bias is when:

1. Studies with significant results are more likely to be published.
2. Studies with non-significant results are more likely to be published.
3. Researchers reported an unexpected finding as having been predicted from the start.

```{r T14}
answer_t14 <- 1
```

## Activity 15

A researcher conducts a study where they compare test scores from native and non-native speakers of English. This is a:

1. Correlational design
2. Within-subjects design
3. Between-subjects design

```{r T15}
answer_t15 <- 3
```

## Activity 16

A researcher conducts a study where they compare reaction times before and after drinking a strong cup of coffee. This is a:

1. Correlational design
2. Within-subjects design
3. Between-subjects design

```{r T16}
answer_t16 <- 2
```

## Activity 17

Which of the following measures of central tendancy is most likely to be influence by the prescence of outliers:

1. Mean
2. Median
3. Mode

```{r T17}
answer_t17 <- 1
```

## Activity 18

Out of the following options, Likert scale data can be best described as:

1. Continuous
2. Categorical
3. Ratio
4. Ordinal

```{r T18}
answer_t18 <- 4
```

## Activity 19

Out of the following options, data on gender can be best described as:

1. Parametric
2. Nominal
3. Interval
4. Skewed

```{r T19}
answer_t19 <- 2
```

## Activity 20

Standard deviation is:

1. Is a measure of the dispersion of sample means around the population mean.
2. A value or quantity lying at the midpoint of a frequency distribution of observed values or quantities, such that there is an equal probability of falling above or below it.
3. On average, how much each data point differs from the mean.
4. The number which appears most often in a set of numbers.

```{r T20}
answer_t20 <- 3
```

## Activity 21

The p-value from an inferential test tells you the probability:

1. Of finding the observed result when the null hypothesis is true
2. That the results were due to chance.
3. Of finding the observed result when the null hypothesis is false. 
4. That there is a significant difference between the IV and the DV.

```{r T21}
answer_t21 <- 1
```

## Activity 22

What is the default cut-off for p-values that psychologists tend to use to assess statistical significance?

1. .5
2. .05
3. .005
4. .0005

```{r T22}
answer_t22 <- 2
```

# Finished

Well done, you are finshed! Please save your code, close Rstudio, reopen Rstudio, reopen your code, and then knit your code one last time. If there are no errors then your code should produce an HTML output with all your answers in place. If any errors appear you should try and rectify them before submitting the .Rmd file to Moodle. Your feedback will be released on Moodle within one week from the submission date. 