---
title: "analysis_template"
date: "20/01/2020"
---

#### 1. Load in packages and data

```{r}
library(tidyverse)
library(dplyr)
library(car)
library(lsr)
library (Hmisc)
library(ggpubr)

demo <- read_csv("demographics.csv")
mslq <- read_csv("mslq.csv")
```

#### 2. Clean up the data

```{r}
demo_final <- demo %>% 
  group_by(user_id, q_id) %>% 
  filter(session_id == min(session_id), endtime == min(endtime)) %>% 
  filter(row_number() == 1) %>% 
  ungroup() %>% 
  filter(user_status %in% c("guest", "registered")) %>%
  select(user_id, user_sex, user_age, q_name, dv) %>%
  pivot_wider(names_from = q_name, values_from = dv)

mslq_final <- mslq %>% 
  group_by(user_id, q_id) %>% 
  filter(session_id == min(session_id), endtime == min(endtime)) %>% 
  filter(row_number() == 1) %>% 
  ungroup() %>% 
  filter(user_status %in% c("guest", "registered")) %>%
  select(user_id, user_sex, user_age, q_name, dv) %>%
  arrange(q_name) %>%
  pivot_wider(names_from = q_name, values_from = dv)

```

#### 3. Join together the data files by their common columns

```{r}
join <- inner_join(mslq_final, demo_final, by = c("user_id", "user_sex", "user_age"))
```

#### 4. Use select to retain only the desired variables.

```{r}
new_dat_anxiety <- select (.data= join,user_id, anxiety_1,anxiety_2,anxiety_3,anxiety_4,anxiety_5, user_sex)

new_dat_efficacy <- select (.data= join,user_id, self_efficacy_1,self_efficacy_2,self_efficacy_3,self_efficacy_4,self_efficacy_5,self_efficacy_6,self_efficacy_7,self_efficacy_8, user_sex)
```

#### 5. Use filter to retain only the observations needed

```{r}
old_dat_anxiety_filtered <- filter (new_dat_anxiety, user_sex=="male"|user_sex=="female")

old_dat_efficacy_filtered <- filter (new_dat_efficacy, user_sex=="male"|user_sex=="female")

new_dat_anxiety_filtered <- drop_na(old_dat_anxiety_filtered)

new_dat_efficacy_filtered <- drop_na(old_dat_efficacy_filtered)
```

#### 6. Calculate the mean score for each participant for each sub-scale. 

```{r eval = TRUE}
temp <- new_dat_anxiety_filtered %>% 
  gather(var, val, anxiety_1:anxiety_5)

dat_means_anxiety <- new_dat_anxiety_filtered %>% 
  gather(var, val, anxiety_1:anxiety_5) %>% 
  group_by_at(vars(-val, -var)) %>% # group by everything except the val and var columns
  summarise(anxiety_mean = mean(val, na.rm = TRUE)) %>% 
  ungroup() 

temp2 <- new_dat_efficacy_filtered %>%
  gather(var, val, self_efficacy_1:self_efficacy_8)
dat_means_efficacy <- new_dat_efficacy_filtered %>% 
  gather(var, val, self_efficacy_1:self_efficacy_8) %>% 
  group_by_at(vars(-val, -var)) %>% 
  summarise(self_efficacy_mean = mean(val, na.rm = TRUE)) %>% 
  ungroup() 
```

#### 7. Data Visualization

T-test visualisation

```{r}
# create summary data to use with `geom_pointrange()`
summary_dat_anxiety<-dat_means_anxiety %>%
  group_by(user_sex)%>%
  summarise(mean = mean(anxiety_mean),
            min = mean(anxiety_mean) - 1.96*sd(anxiety_mean)/sqrt(n()), # 95% confidence intervals
            max = mean(anxiety_mean) + 1.96*sd(anxiety_mean)/sqrt(n()))


ggplot(dat_means_anxiety, aes(x = user_sex, y =anxiety_mean )) +
  geom_violin(trim = FALSE) +
  geom_boxplot(aes(fill = user_sex), width = .2, show.legend = FALSE) + 
  geom_pointrange(data = summary_dat_anxiety,
                  aes(x = user_sex, y = mean, ymin=min, ymax=max),
                  shape = 20, 
                  position = position_dodge(width = 0.1), show.legend = FALSE)

ggplot(dat_means_anxiety, aes(x = user_sex, y = anxiety_mean, fill=user_sex))+
  geom_boxplot()+
  geom_jitter(height = 0, width = .1)


summary_dat_efficacy<-dat_means_efficacy%>%
  group_by(user_sex)%>%
  summarise(mean = mean(self_efficacy_mean),
            min = mean(self_efficacy_mean) - 1.96*sd(self_efficacy_mean)/sqrt(n()), # 95% confidence intervals
            max = mean(self_efficacy_mean) + 1.96*sd(self_efficacy_mean)/sqrt(n()))


ggplot(dat_means_efficacy, aes(x = user_sex, y = self_efficacy_mean )) +
  geom_violin(trim = FALSE) +
  geom_boxplot(aes(fill = user_sex), width = .2, show.legend = FALSE) + 
  geom_pointrange(data = summary_dat_efficacy,
                  aes(x = user_sex, y = mean, ymin=min, ymax=max),
                  shape = 20, 
                  position = position_dodge(width = 0.1), show.legend = FALSE)

ggplot(dat_means_efficacy, aes(x = user_sex, y = self_efficacy_mean, fill=user_sex))+
  geom_boxplot()+
  geom_jitter(height = 0, width = .1)
```



#### 8. Check that the data meets the assumptions of t-test and correlation

Correlation visualization and checking t-test assumptions

```{r}

qq_anxiety_male <- filter(dat_means_anxiety, user_sex=="male") %>%
  group_by(user_sex) %>%
  mutate(group_resid = anxiety_mean - mean(anxiety_mean)) %>%
  ungroup()

ggdensity(qq_anxiety_male$group_resid)
qqPlot(qq_anxiety_male$group_resid)

shapiro.test(x = qq_anxiety_male$group_resid)

qq_anxiety_female <- filter(dat_means_anxiety, user_sex=="female") %>%
  group_by(user_sex) %>%
  mutate(group_resid = anxiety_mean - mean(anxiety_mean)) %>%
  ungroup()

ggdensity(qq_anxiety_female$group_resid)
qqPlot(qq_anxiety_female$group_resid)

shapiro.test(x = qq_anxiety_female$group_resid[0:5000]) 

qq_self_efficacy_male <- filter(dat_means_efficacy, user_sex=="male") %>%
  group_by(user_sex) %>%
  mutate(group_resid = self_efficacy_mean - mean(self_efficacy_mean)) %>%
  ungroup()

ggdensity(qq_self_efficacy_male$group_resid)
qqPlot(qq_self_efficacy_male$group_resid)

shapiro.test(x = qq_self_efficacy_male$group_resid)

qq_self_efficacy_female <- filter(dat_means_efficacy, user_sex=="female") %>%
  group_by(user_sex) %>%
  mutate(group_resid = self_efficacy_mean - mean(self_efficacy_mean)) %>%
  ungroup()

ggdensity(qq_self_efficacy_female$group_resid)
qqPlot(qq_self_efficacy_female$group_resid)

shapiro.test(x = qq_self_efficacy_female$group_resid[0:5000])

```

Check Correlation assumptions

```{r}

cor_visual<- inner_join(dat_means_anxiety,dat_means_efficacy, by = c("user_id", "user_sex"))

ggplot(data = cor_visual, aes(x = self_efficacy_mean, y = anxiety_mean, colour= user_sex)) +
  geom_jitter() +
  scale_x_continuous(name= "Self-Efficacy")+
scale_y_continuous(name= "Test Anxiety")+
  geom_smooth(method = lm)

```


#### 9. Conduct statistical analyses

T-test analysis

```{r}
t.test(dat_means_anxiety$anxiety_mean~ dat_means_anxiety$user_sex, alternative= "greater")

t.test(dat_means_efficacy$self_efficacy_mean~ dat_means_efficacy$user_sex, alternative= "less")
```

Correlation analysis

```{r}
cor.test(cor_visual$self_efficacy_mean, cor_visual$anxiety_mean, method="pearson", alternative= "two.sided" )
```



